---
title: "project_name"
author: "your_name"
output:
    html_document: default
    html_notebook: default
---

# Part 5: Marker genes

```{r setup_opt, warning=FALSE,error=FALSE,message=FALSE, echo = F, eval=F}
options(width = 450)
```

## Load libraries
```{r load_libraries, warning=FALSE,error=FALSE,message=FALSE}
library(Seurat)
library(ggplot2)
library(dplyr)
library(future)
```

```{r}
run_tests = F
```


## Load the Seurat object
```{r user_vals, warning=FALSE,error=FALSE,message=FALSE, eval=!run_tests}
r.obj.loc <- "r_obj_path"

# Set this to the resolution that you want to do downstream analysis on
Idents(experiment.merged) <- "RNA_snn_res.0.25"
num.cores.available = 1 # set to a different number to get multithreading when running findAllMarkers
```


```{r load_rdata_test, warning=FALSE,error=FALSE,message=FALSE, eval= run_tests}
r.obj.loc <- "/Volumes/My_Passport_for_Mac/tenX_projects/Bonini05/iAS_Pb_project/rds/all_samples/"
Idents(experiment.aggregate) <- "RNA_snn_res.0.25"
```

```{r load_rdata, warning=FALSE,error=FALSE,message=FALSE}
load(file=paste0(r.obj.loc,"clusters_seurat_object.RData"))
```

### Choosing a resolution

Table of cluster to sample assignments.
```{r cells_per_res_sample, warning=FALSE,error=FALSE,message=FALSE}
table(Idents(experiment.aggregate),experiment.aggregate$orig.ident)
```

<!-- Plot TSNE coloring by the slot 'ident' (default). -->
<!-- ```{r plot_tsne, warning=FALSE,error=FALSE,message=FALSE} -->
<!-- DimPlot(object = experiment.aggregate, pt.size=0.5, reduction = "tsne", label = T) -->
<!-- ``` -->


### Plot uMap coloring by the slot 'ident' (default).
```{r plot_umap, warning=FALSE,error=FALSE,message=FALSE}
DimPlot(object = experiment.aggregate, pt.size=0.5, reduction = "umap", label = T)
```

### Plot UMAP  coloring by the slot 'orig.ident' (sample names) with alpha colors turned on. A pretty picture
```{r pretty_pre, warning=FALSE,error=FALSE,message=FALSE}
DimPlot(object = experiment.aggregate, group.by="orig.ident", pt.size=0.5, reduction = "umap" )
```



### uMAP plot by cell cycle
```{r plot_cellcycle, warning=FALSE,error=FALSE,message=FALSE}
DimPlot(object = experiment.aggregate, pt.size=0.5, group.by = "Phase", reduction = "umap" )
```


### Use feature plot to plot our read valued metadata, like nUMI, Feature count, and percent Mito

FeaturePlot can be used to color cells with a 'feature', non categorical data, like number of UMIs
```{r plot_rna, warning=FALSE,error=FALSE,message=FALSE}
FeaturePlot(experiment.aggregate, features = c('nCount_RNA'), pt.size=0.5)
```
and number of genes present
```{r plot_feature, warning=FALSE,error=FALSE,message=FALSE}
FeaturePlot(experiment.aggregate, features = c('nFeature_RNA'), pt.size=0.5)
```

percent mitochondrial
```{r plot_mito, warning=FALSE,error=FALSE,message=FALSE}
FeaturePlot(experiment.aggregate, features = c('percent.mito'), pt.size=0.5)
```


## Identifying Marker Genes

Seurat can help you find markers that define clusters via differential expression.

`FindMarkers` identifies markers for a cluster relative to all other clusters.

`FindAllMarkers` does so for all clusters

`FindAllMarkersNode` defines all markers that split a Node from the cluster tree

### FindAllMarkers can be used to automate the process across all genes.

```{r find_all_markers, warning=FALSE,error=FALSE,message=FALSE, eval=FALSE}
# DefaultAssay(experiment.aggregate) <- 'RNA'

plan("multiprocess", workers = num.cores.available) # multithreading

markers_all <- FindAllMarkers(
    object = experiment.aggregate,
    only.pos = TRUE,
    min.pct = 0.25,
    thresh.use = 0.25
)
```


pct.1 and pct.2 are the proportion of cells with expression above 0 in ident.1 and ident.2 respectively. p_val is the raw p_value associated with the differntial expression test with adjusted value in p_val_adj. avg_logFC is the average log fold change difference between the two groups.

avg_diff appears to be the difference in log(x = mean(x = exp(x = x) - 1) + 1) between groups.  It doesn’t seem like this should work out to be the signed ratio of pct.1 to pct.2 so I must be missing something.  It doesn’t seem to be related at all to how the p-values are calculated so maybe it doesn’t matter so much, and the sign is probably going to be pretty robust to how expression is measured.


### Feature plots and Violin Plots of Marker Genes
```{r gene_feature, warning=FALSE,error=FALSE,message=FALSE}
feat_plots_top_genes <- function(cluster.num){  
  print(paste0("Cluster number ", cluster.num))
  curr.markers <- markers_all[markers_all$cluster == cluster.num,] 
  # print(head(curr.markers))
  print(FeaturePlot( 
    experiment.aggregate,
    curr.markers$gene[1:6],
    cols = c("lightgrey", "blue"),
    ncol = 3 ))
  
  print(VlnPlot(object = experiment.aggregate, 
                features = curr.markers$gene[1:6], 
                pt.size = 0.05))
}

# these.markers <- markers_all[markers_all$cluster == 6,]
# VlnPlot(object = experiment.aggregate, features = these.markers$gene[1:6], pt.size = 0.05)
# num.clusters.list = c(0, 1, 2)
num.clusters.list <- unique(markers_all$cluster)
# num.clusters.list
sapply(num.clusters.list, feat_plots_top_genes)
```

<br/>
<br/>

Plot a heatmap of genes by cluster for the top 10 marker genes per cluster
```{r markers_head, warning=FALSE,error=FALSE,message=FALSE, fig.height=10}
top10 <- markers_all_single %>% group_by(cluster) %>% top_n(10, avg_log2FC)
DoHeatmap(
    object = experiment.aggregate,
    features = top10$gene
)
```

And last lets save all the objects in our session.
```{r save_rdata, warning=FALSE,error=FALSE,message=FALSE, eval=FALSE}
# We want to save everything in the environment, especially the seurat object and the list of marker genes, so use ls()
save(list=ls(), file=paste0(r.obj.loc,"clustered_seurat_object_and_markers.RData"))
# save(markers_all, file="clusters_seurat_markers_all.RData")
# save(experiment.aggregate, file=paste0(r.obj.loc, "clusters_seurat_object.RData"))
```

## Session Information
```{r session_info, warning=FALSE,error=FALSE,message=FALSE}
sessionInfo()
```
